---
title: "Mortality Rates and Causes Across the United States using"
author: "Subhalaxmi Rout"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: architect
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

### Instruction

I have provided you with data about mortality from all 50 states and the District of Columbia.
Please access it at
https://github.com/charleyferrari/CUNY_DATA608/tree/master/lecture3/data

**Question 1:**
As a researcher, you frequently compare mortality rates from particular causes across
different States. You need a visualization that will let you see (for 2010 only) the crude
mortality rate, across all States, from one cause (for example, Neoplasms, which are
effectively cancers). Create a visualization that allows you to rank States by crude mortality
for each cause of death.

### Load libraries

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(stats)
library(DT)
```

### Load data from Git

```{r}
df <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv", header = TRUE)

head(df)
```

The below plot shows state by crude rate filtered with year = 2010, and cause for death = Neoplasms. 

```{r, fig.height=10, fig.width=10}
df2 <- df %>% filter(Year == '2010' & ICD.Chapter == 'Neoplasms') %>% select(State, Crude.Rate)

df2 %>% group_by(State) %>% summarise(Crude.Rate = sum(Crude.Rate)) %>%
  ggplot() + aes(x = reorder(State, Crude.Rate), y = Crude.Rate) +
  ggtitle('States wise Mortality Rate') +
  xlab('State') +
  geom_bar(fill="#276678", stat = "identity") +
  coord_flip() + geom_text(aes(label = Crude.Rate), size = 5,  hjust=-0.20) +
  theme(panel.background = element_rect(fill = "white", color = NA),
        plot.title = element_text(hjust = 0.5, size = 25, colour = "#03506f"),
        axis.title.y = element_text(size = 20, colour = "#03506f"),
        axis.title.x=element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x=element_blank()
       )
```

Question 2:
Often you are asked whether particular States are improving their mortality rates (per cause)
faster than, or slower than, the national average. Create a visualization that lets your clients
see this for themselves for one cause of death at the time. Keep in mind that the national
average should be weighted by the national population.

```{r}

df3 <- df  %>%  filter(ICD.Chapter == 'External causes of morbidity and mortality') %>% group_by(Year) %>%
  mutate(National.Crude.Rate = sum(Deaths) / sum(Population) * 100000) 
  
df4 <- df %>% filter(ICD.Chapter == 'External causes of morbidity and mortality') %>% group_by(Year)

df5 <- cbind(df4, df3$National.Crude.Rate)

df5 <- df5 %>% filter(State == 'AK')

colnames(df5)[7] <- "National.Crude.Rate" 

df5$National.Crude.Rate = round(df5$National.Crude.Rate,1)
DT::datatable(df5)
```

```{r}

df5 <- df5 %>% select(Year, Crude.Rate, National.Crude.Rate) %>% gather("Crude", "Value", -Year)
head(df5)
```

The below plot shows Yearly Crude Rate vs National Crude Rate filtered with State = AK, and cause for death = Certain infectious and parasitic diseases. 

```{r, fig.height=10, fig.width=10}
df5 %>% 
  ggplot() + aes(x = Year, y = Value, fill = Crude) +
  geom_col(position = "dodge") +
  ggtitle('Crude Rate vs National Crude Rate') +
  xlab('Year') +
  ylab('Crude Rate') +
  scale_fill_manual(values = c("#276678", "#999999")) +
  geom_text(aes(label = Value), size = 4, position = position_dodge(width = 1), vjust = -.50) +
  theme(panel.background = element_rect(fill = "white", color = NA),
        plot.title = element_text(hjust = 0.5, size = 15, colour = "#03506f"),
        axis.title.y = element_text(size = 10, colour = "#03506f")
       )
```

